---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# paar

<!-- badges: start -->
[![R-CMD-check](https://github.com/PPaccioretti/paar/workflows/R-CMD-check/badge.svg)](https://github.com/PPaccioretti/paar/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of `paar` is to provide useful functions for 
precision agriculture spatial data depuration. 

## Installation

<!-- You can install the released version of paar from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("paar") -->
<!-- ``` -->
You can install the development version from [GitHub](https://github.com/PPaccioretti/paar) with:

``` r
# install.packages("devtools")
devtools::install_github("PPaccioretti/paar")
```

## Example

The package has a complete protocol for automating error removal. Default
values of all functions are optimized for precision agricultural data.

```{r example, message=FALSE}
library(paar)
library(sf)
data("barley", package = 'paar')
```

`barley` data contains barley grain yield which were obtained using calibrated commercial yield monitors, mounted on combines equipped with DGPS.

```{r barley}
#Convert barley data to an spatial object
barley_sf <- st_as_sf(barley,
                   coords = c("X", "Y"),
                   crs = 32720)

barley_dep <-
  depurate(barley_sf,
           "Yield")

# Summary of depurated data
summary(barley_dep)
```

Spatial yield values before and after depuration process can be plotted

```{r, fig.height=2, fig.height=5, out.width="50%", fig.show="hold"}
plot(barley_sf["Yield"], main = "Before depuration")
plot(barley_dep$depurated_data["Yield"], main = "After depuration")
```

Also distribution of yield values can be plotted

```{r, fig.height=2, fig.height=5, out.width="50%", fig.show="hold"}
boxplot(barley_sf[["Yield"]], main = "Before depuration")
boxplot(barley_dep$depurated_data[["Yield"]], main = "After depuration")
```


## References

- Vega, A., Córdoba, M., Castro-Franco, M. et al. (2019). Protocol for automating error removal from yield maps. Precision Agric 20, 1030–1044  https://doi.org/10.1007/s11119-018-09632-8
- Paccioretti, P., Córdoba, M., & Balzarini, M. (2020). FastMapping: Software to create field maps and identify management zones in precision agriculture. Computers and Electronics in Agriculture, 175, 105556 https://doi.org/10.1016/j.compag.2020.105556.


