---
title: "Yield data preprocessing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Yield_data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Usually, yield data comes with many noisy observations. This vignette
will show how to preprocess yield data to remove both, spatial and
global outliers. The protocol for error removal follows the protocol
proposed by [@Vega2019]. For the tutorial we
will use the `barley` dataset that comes with the `paar` package. The
`barley` data contains barley grain yield which were obtained using
calibrated commercial yield monitors, mounted on combines equipped with
DGPS. The data is not a `sf` object format. We will convert it to an
`sf` object first.

First, we will load the `paar` package, the `sf` package for spatial data manipulation, `ggplot2` for plotting, and the `barley` dataset that comes with 
the `paar` package.

```{r setup}
library(paar)
library(sf)
library(ggplot2)

data("barley", package = 'paar')
```

The `barley` dataset is a `data.frame` object. We will convert it to a
`sf` object using the `st_as_sf` function. The `coords` argument
specifies the columns that contain the coordinates. The `crs` argument
specifies the coordinate reference system. The `barley` dataset is in
UTM zone 20S.

```{r}
barley_sf <- st_as_sf(barley, 
                      coords = c("X", "Y"),
                      crs = 32720)
```

The `barley_sf` object is now an `sf` object. We can plot the data to
visualize the yield data.

```{r, fig.height=2, fig.height=5, out.width="50%", fig.show="hold"}
plot(barley_sf["Yield"])

ggplot(barley_sf) +
  geom_sf(aes(fill = Yield)) +
  scale_fill_viridis_c() +
  theme_minimal()
```

Let's see the yield values distribution.

```{r, fig.height=2, fig.height=5, out.width="50%", fig.show="hold"}
hist(barley_sf$Yield)

ggplot(barley_sf) +
  geom_histogram(aes(x = Yield)) +
  theme_minimal()
```


The protocol proposed by [@Vega2019], is implemented in the function `depurate`
and consists of three steps:
1. Remove border observations (_edges_).
2. Remove global outliers (_outliers_).
3. Remove spatial outliers (_inliers_).

The `depurate` function takes an `sf` object as input and returns an object
of class `paar`. Any combination of 
the three steps can be done using the `depurate` function. The argument
`to_remove` specifies which steps to perform. The argument `y` specifies the
column name of the variable to be cleaned. 
A field boundary is necessary to remove the _edges_ observations. If a polygon
is not provided in the `poly_border` argument, the function will make a hull,
around the data and remove the observation that are 10m from the hull.
The hull is made using `concaveman::concaveman` function if the package 
is installed, otherwise, the `sf::st_convex_hull` function is used.


```{r}
barley_clean <- depurate(barley_sf, 
                         y = 'Yield',
                         toremove = c("edges", "outlier", "inlier"))


```


The `depurate` function returns an object of class `paar`. The
`paar` object contains the cleaned data (`$depurated_data`), and the condition 
of each observation (`$condition`). If the condition is `NA` means that 
the observation was not removed.

```{r}
barley_clean
```

The `summary` function can be used to get a summary of the percentage of
considered ouliers and the number of observations removed.


```{r}
summary_table <- summary(barley_clean)
summary_table
```

The condition of each observation can be combined to the original data using the 
`cbind` function.


```{r}
barley_sf <- cbind(condition = barley_clean$condition, barley_sf)

```





