---
title: "Zone delineation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Zone delineation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
has_ggplot <- requireNamespace("ggplot2")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(paar)
library(sf)
require(ggplot2)
```

Multivariate zone delineation can be done using `kmspc` function, whereas 
univariate zone delineation can be done with `fyzzy_k_means` function.

Multivariate function implements the protocol proposed by @Cordoba2016, which
performs a clustering with the `kmeans` function using as an input the spatial
principal components (sPC) of the data. 
The function requires an `sf` object with the data to be clustered, and more than
one numeric variable. The function by default returns a list with the 
following components:
- `summaryResults`: a `data.frame` with 
- `indices`: a `data.frame` with indices to help to chose the optimal number 
of clusters.
- `cluster`: the cluster number assigned to each observation.

For this example we will use the `wheat` dataset that comes with the `paar` package.
The `data.frame` has apparent electrical conductivity (ECa) meassured at 
two depths, elevation data, soil depth, and wheat gran yield. All variables 
have been interpolated to an unique grid and then merged in a single `data.frame`.


```{r}
data(wheat, package = 'paar')

wheat_sf <- st_as_sf(wheat,
                     coords = c('x', 'y'),
                     crs = 32720)
```

```{r}
plot(wheat_sf)
```

The function `kmspc` requires the `sf` object with the data to be clustered, and
the number of clusters (zones) to be delineated.
For the sPC process, is necessary to specify the distance in wshich observations 
will be considered neighbors. The `ldist` and `udist` arguments specify the
lower and upper distance thresholds, respectively. The `explainedVariance`
argument specifies the minimum value of explained variance that the Principal 
Component to be used for the cluster process should explain. The `center` 
argument specifies if the data should be centered before the sPC process 
(default `TRUE`).

```{r}
# Run the kmspc function
kmspc_results <- 
  kmspc(wheat_sf,
        number_cluster = 2:4,
        explainedVariance = 70,
        ldist = 0,
        udist = 40,
        center = TRUE)

```

To help the user to chose the optimal number of clusters, the function returns
a `data.frame` with indices (`Xie Beni`, `Partition Coefficient`, 
`Entropy of Partition`, and `Summary Index`).
The `Summary Index` is a combination of the indices to obtain a measure of the
information reported by each index. In this example, the optimuin number of
cluster is 2. For each index, lower the value better the
clustering. More information can be found in @Paccioretti2020.


```{r}
kmspc_results$indices
```


The cluster for each observation can be found in the `cluster` component of the
`kmspc` object. The `cluster` component is a `data.frame` with the cluster
number assigned to each observation.

```{r}
head(kmspc_results$cluster)
```

The clusters can be combined to the original data using the `cbind` function.

```{r}
wheat_clustered <- cbind(wheat_sf, kmspc_results$cluster)
```


This cluster can be plotted with the `plot` function.

```{r}
plot(wheat_clustered[, "Cluster_2"])
```


Also, ggplot can be used to plot the clusters.
```{r eval = !has_ggplot, echo = FALSE, comment = NA}
message("No package ggplot2 available. Code chunks using that package will not be evaluated.")
```

```{r, eval = has_ggplot}
ggplot(wheat_clustered) +
  geom_sf(aes(color = Cluster_2)) +
  theme_minimal()
```

